---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---



## Loading and preprocessing the data


```{r}
fileUrl <- "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"
if (!file.exists("activity.csv")) {
  download.file(fileUrl, destfile="activity.zip", method="curl")
  unzip("activity.zip")
}
data <- read.csv("activity.csv")
```



## What is mean total number of steps taken per day?


1. Calculate the total number of steps taken per day  

We will use the dplyr librarie : `library(dplyr)`  
We calculate the total of steps after grouping by day
```{r echo=FALSE, results='hide'}
library(dplyr)
```

```{r}
per_day <- group_by(data, date)
step_per_day <- summarise(per_day, total_step = sum(steps, na.rm = TRUE))
step_per_day
```


2. Make a histogram of the total number of steps taken each day  

```{r}
hist(step_per_day$total_step, col = "blue", main = "Total number of steps per day",
     xlab = "number of steps")
```


3. Calculate and report the mean and median of the total number of steps taken per day  

```{r}
mean_steps_per_day <- mean(step_per_day$total_step, na.rm = TRUE)
mean_steps_per_day
median_steps_per_day <- median(step_per_day$total_step, na.rm = TRUE)
median_steps_per_day
```


## What is the average daily activity pattern?  

1. Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis)
and the average number of steps taken, averaged across all days (y-axis)  

```{r}
interval_per_day <- summarise(group_by(per_day, interval), 
                              avg_steps = mean(steps, na.rm = TRUE))

plot(interval_per_day$interval, interval_per_day$avg_steps, type = "l",
     main = "average daily activity", xlab = "5-minute interval",
     ylab = "average number of steps")
```

2. Which 5-minute interval, on average across all 
the days in the dataset, contains the maximum number of steps?  

```{r}
max_steps <- filter(interval_per_day, avg_steps == max(interval_per_day$avg_steps, na.rm = TRUE) )
max_steps$interval
```



## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
